---
output:
  html_document:
    code_folding: "hide"
    toc: true
    toc_float: true
---
<section class="page-header">
<h1 class="title toc-ignore project-name" style="text-align:center;">Marketing Data Analysis</h1>
<h4 class="author project-author" style="text-align:center;">Hanie Jalili</h4>
<h4 class="date project-date" style="text-align:center;">2022-11-10</h4>
<body style="background-color:aliceblue;">
</section>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

The data is related to a chain company. The purpose of collecting this data is to check the impact of this company's advertising campaigns on its customers.

In this data, we have 26 variables that are related to the demographic information of customers, customer behavior, the effectiveness of advertisements, and the amount of purchase of various products by customers.

## Exploring and cleaning the data set

### Exploring

First, we start by loading the data into R Studio to view it.

```{r echo=TRUE}
data<-read.table("E:\\Users\\ASUS\\Documents\\term7\\multivariate analysis\\mydata.csv",header = T,sep = ",")
attach(data)
dim(data)
colnames(data)
```

We have 2240 observations and 26 variables. Now we summarize the dataset.

```{r echo=TRUE}
library(skimr)
skim(data)
```

### Cleaning the dataset

For data analysis, it is necessary to change the data type to numerical and factorial type.

```{r echo=TRUE}
Birth <- as.numeric((Birth))
Income <- as.numeric((Income))
Recency <- as.numeric((Recency))
Wines <- as.numeric((Wines))
Fruits <- as.numeric((Fruits))
Meat <- as.numeric((Meat))
Fish <- as.numeric((Fish))
Sweet <- as.numeric((Sweet))
Gold <- as.numeric((Gold))
Deals <- as.numeric((Deals))
Web <- as.numeric((Web))
Catalog <- as.numeric((Catalog))
Store <- as.numeric((Store))
WebVisitsMonth <- as.numeric((WebVisitsMonth))

Education <- as.factor((Education))
Marital_Status <- as.factor((Marital_Status))
Kid <- as.factor((Kid))
Teen <- as.factor((Teen))
Complain <- as.factor((Complain))
AcceptedCmp3 <- as.factor((AcceptedCmp3))
AcceptedCmp4 <- as.factor((AcceptedCmp4))
AcceptedCmp5 <- as.factor((AcceptedCmp5))
AcceptedCmp1 <- as.factor((AcceptedCmp1))
AcceptedCmp2 <- as.factor((AcceptedCmp2))
Response <- as.factor((Response))
```

In the previous investigations, we saw that only the "income" variable has missing values. To find the best replacement value for the missing values, we first pay attention to the distribution of the "income" variable:

```{r echo=TRUE}
library(ggplot2)
  ggplot(data, aes(x=Income)) +
  geom_density(fill="#69b3a2", color="#e9ecef", alpha=0.8)
```

```{r echo=TRUE}
shapiro.test(Income)
```

The assumption of normality of income distribution was rejected and we also see that its distribution is not symmetrical.
So it is better to replace the missing values with the median of the income variable.

```{r echo=TRUE}
sum(is.na(Income))
m = median(Income,na.rm = TRUE) 
sel = is.na(Income)
Income[sel] = m
```

To check outliers in quantitative data, we divide them into the following three categories:

* The number of purchases made in different ways
* Consumption of different products in the last two years
* Income

```{r echo=TRUE}
num_purchases=data.frame(Deals,Web,Catalog,Store,WebVisitsMonth)
income=data.frame(Income)
mnt_products=data.frame(Wines,Fruits,Meat,Fish,Sweet,Gold)
summary(num_purchases)
summary(Income)
summary(mnt_products)
```

Now, to see the outlier values more clearly, we draw their graphs:

```{r echo=TRUE}
boxplot(num_purchases, col = rainbow(ncol(num_purchases)))
boxplot(income, col = rainbow(ncol(income)))
boxplot(mnt_products, col = rainbow(ncol(mnt_products)))
```

As can be seen in the graphs, there are many outliers in each variable. Therefore, we cannot remove all of them from the model at the same time.
For this reason, I just arbitrarily selected one of the variables and removed its outliers from the dataset.My chosen variable is income.After removing outliers of "income", we create a new dataset that name is "new_data".

```{r echo=TRUE}
new_data<- filter(data,Income <= 200000)
new_data=as.data.frame(new_data)
```

## Analysis & results

Our goal of analyzing this data is to investigate the effect of different levels of the advertising campaign on other variables.

### The effect of campaigns on consumption of products

```{r out.width=c('50%', '50%'), fig.show='hold'}
library(ggplot2)
ggplot(data=new_data, aes(x=AcceptedCmp1, y=Gold)) +
  geom_bar(stat="identity", fill="pink")
ggplot(new_data, aes(x=AcceptedCmp2, y=Gold)) + 
  geom_bar(stat = "identity", fill="pink")
ggplot(new_data, aes(x=AcceptedCmp3, y=Gold)) + 
  geom_bar(stat = "identity", fill="pink")
ggplot(new_data, aes(x=AcceptedCmp4, y=Gold)) + 
  geom_bar(stat = "identity", fill="pink")
ggplot(new_data, aes(x=AcceptedCmp5, y=Gold)) + 
  geom_bar(stat = "identity", fill="pink")
ggplot(new_data, aes(x=Response, y=Gold)) + 
  geom_bar(stat = "identity", fill="pink")
```

The last advertising campaign had the greatest impact on gold purchases in the last two years, and the second campaign had the least impact on gold purchases.

```{r out.width=c('50%', '50%'), fig.show='hold'}
library(ggplot2)
ggplot(new_data, aes(x=AcceptedCmp1, y=Meat)) + 
  geom_bar(stat = "identity", fill="seagreen3")
ggplot(new_data, aes(x=AcceptedCmp2, y=Meat)) + 
  geom_bar(stat = "identity", fill="seagreen3")
ggplot(new_data, aes(x=AcceptedCmp3, y=Meat)) + 
  geom_bar(stat = "identity", fill="seagreen3")
ggplot(new_data, aes(x=AcceptedCmp4, y=Meat)) + 
  geom_bar(stat = "identity", fill="seagreen3")
ggplot(new_data, aes(x=AcceptedCmp5, y=Meat)) + 
  geom_bar(stat = "identity", fill="seagreen3")
ggplot(new_data, aes(x=Response, y=Meat)) + 
  geom_bar(stat = "identity", fill="seagreen3")
```

```{r out.width=c('50%', '50%'), fig.show='hold'}
library(ggplot2)
ggplot(new_data, aes(x=AcceptedCmp1, y=Sweet)) + 
  geom_bar(stat = "identity", fill="turquoise4")
ggplot(new_data, aes(x=AcceptedCmp2, y=Sweet)) + 
  geom_bar(stat = "identity", fill="turquoise4")
ggplot(new_data, aes(x=AcceptedCmp3, y=Sweet)) + 
  geom_bar(stat = "identity", fill="turquoise4")
ggplot(new_data, aes(x=AcceptedCmp4, y=Sweet)) + 
  geom_bar(stat = "identity", fill="turquoise4")
ggplot(new_data, aes(x=AcceptedCmp5, y=Sweet)) + 
  geom_bar(stat = "identity", fill="turquoise4")
ggplot(new_data, aes(x=Response, y=Sweet)) + 
  geom_bar(stat = "identity", fill="turquoise4")
```

```{r out.width=c('50%', '50%'), fig.show='hold'}
library(ggplot2)
ggplot(new_data, aes(x=AcceptedCmp1, y=Fish)) + 
  geom_bar(stat = "identity", fill="mediumvioletred")
ggplot(new_data, aes(x=AcceptedCmp2, y=Fish)) + 
  geom_bar(stat = "identity", fill="mediumvioletred")
ggplot(new_data, aes(x=AcceptedCmp3, y=Fish)) + 
  geom_bar(stat = "identity", fill="mediumvioletred")
ggplot(new_data, aes(x=AcceptedCmp4, y=Fish)) + 
  geom_bar(stat = "identity", fill="mediumvioletred")
ggplot(new_data, aes(x=AcceptedCmp5, y=Fish)) + 
  geom_bar(stat = "identity", fill="mediumvioletred")
ggplot(new_data, aes(x=Response, y=Fish)) + 
  geom_bar(stat = "identity", fill="mediumvioletred")
```

```{r out.width=c('50%', '50%'), fig.show='hold'}
library(ggplot2)
ggplot(new_data, aes(x=AcceptedCmp1, y=Fruits)) + 
  geom_bar(stat = "identity", fill="orangered")
ggplot(new_data, aes(x=AcceptedCmp2, y=Fruits)) + 
  geom_bar(stat = "identity", fill="orangered")
ggplot(new_data, aes(x=AcceptedCmp3, y=Fruits)) + 
  geom_bar(stat = "identity", fill="orangered")
ggplot(new_data, aes(x=AcceptedCmp4, y=Fruits)) + 
  geom_bar(stat = "identity", fill="orangered")
ggplot(new_data, aes(x=AcceptedCmp5, y=Fruits)) + 
  geom_bar(stat = "identity", fill="orangered")
ggplot(new_data, aes(x=Response, y=Fruits)) + 
  geom_bar(stat = "identity", fill="orangered")
```

```{r out.width=c('50%', '50%'), fig.show='hold'}
library(ggplot2)
ggplot(new_data, aes(x=AcceptedCmp1, y=Wines)) + 
  geom_bar(stat = "identity", fill="palegoldenrod")
ggplot(new_data, aes(x=AcceptedCmp2, y=Wines)) + 
  geom_bar(stat = "identity", fill="palegoldenrod")
ggplot(new_data, aes(x=AcceptedCmp3, y=Wines)) + 
  geom_bar(stat = "identity", fill="palegoldenrod")
ggplot(new_data, aes(x=AcceptedCmp4, y=Wines)) + 
  geom_bar(stat = "identity", fill="palegoldenrod")
ggplot(new_data, aes(x=AcceptedCmp5, y=Wines)) + 
  geom_bar(stat = "identity", fill="palegoldenrod")
ggplot(new_data, aes(x=Response, y=Wines)) + 
  geom_bar(stat = "identity", fill="palegoldenrod")
```

By looking at other charts, we realize that other products behave like gold against advertising campaigns. It means that the last campaign had the most impact on them and the second campaign had the least impact.

### The effect of campaigns on number of purchases

```{r out.width=c('50%', '50%'), fig.show='hold'}
library(ggplot2)
ggplot(new_data, aes(x=AcceptedCmp1, y=Deals)) + 
    geom_boxplot(color="red", fill="orange", alpha=0.2)
ggplot(new_data, aes(x=AcceptedCmp2, y=Deals)) + 
    geom_boxplot(color="red", fill="orange", alpha=0.2)
ggplot(new_data, aes(x=AcceptedCmp3, y=Deals)) + 
    geom_boxplot(color="red", fill="orange", alpha=0.2)
ggplot(new_data, aes(x=AcceptedCmp4, y=Deals)) + 
    geom_boxplot(color="red", fill="orange", alpha=0.2)
ggplot(new_data, aes(x=AcceptedCmp5, y=Deals)) + 
    geom_boxplot(color="red", fill="orange", alpha=0.2)
ggplot(new_data, aes(x=Response, y=Deals)) + 
    geom_boxplot(color="red", fill="orange", alpha=0.2)
```

In campaigns 3 and 4, the distribution of rejecters and accepters are the same, So it means that the effect of them on people are the same.

```{r out.width=c('50%', '50%'), fig.show='hold'}
library(ggplot2)
ggplot(new_data, aes(x=AcceptedCmp1, y=Web)) + 
    geom_boxplot(color="cadetblue", fill="cadetblue1", alpha=0.2)
ggplot(new_data, aes(x=AcceptedCmp2, y=Web)) + 
    geom_boxplot(color="cadetblue", fill="cadetblue1", alpha=0.2)
ggplot(new_data, aes(x=AcceptedCmp3, y=Web)) + 
    geom_boxplot(color="cadetblue", fill="cadetblue1", alpha=0.2)
ggplot(new_data, aes(x=AcceptedCmp4, y=Web)) + 
    geom_boxplot(color="cadetblue", fill="cadetblue1", alpha=0.2)
ggplot(new_data, aes(x=AcceptedCmp5, y=Web)) + 
    geom_boxplot(color="cadetblue", fill="cadetblue1", alpha=0.2)
ggplot(new_data, aes(x=Response, y=Web)) + 
    geom_boxplot(color="cadetblue", fill="cadetblue1", alpha=0.2)
```

In campaigns 1, 4 and 5, the distribution of rejecters and accepters are the same, So it means that they affect people in the same way.

```{r out.width=c('50%', '50%'), fig.show='hold'}
library(ggplot2)
ggplot(new_data, aes(x=AcceptedCmp1, y=Catalog)) + 
    geom_boxplot(color="aquamarine3", fill="aquamarine", alpha=0.2)
ggplot(new_data, aes(x=AcceptedCmp2, y=Catalog)) + 
    geom_boxplot(color="aquamarine3", fill="aquamarine", alpha=0.2)
ggplot(new_data, aes(x=AcceptedCmp3, y=Catalog)) + 
    geom_boxplot(color="aquamarine3", fill="aquamarine", alpha=0.2)
ggplot(new_data, aes(x=AcceptedCmp4, y=Catalog)) + 
    geom_boxplot(color="aquamarine3", fill="aquamarine", alpha=0.2)
ggplot(new_data, aes(x=AcceptedCmp5, y=Catalog)) + 
    geom_boxplot(color="aquamarine3", fill="aquamarine", alpha=0.2)
ggplot(new_data, aes(x=Response, y=Catalog)) + 
    geom_boxplot(color="aquamarine3", fill="aquamarine", alpha=0.2)
```

```{r out.width=c('50%', '50%'), fig.show='hold'}
library(ggplot2)
ggplot(new_data, aes(x=AcceptedCmp1, y=Store)) + 
    geom_boxplot(color="mediumorchid4", fill="mediumorchid1", alpha=0.2)
ggplot(new_data, aes(x=AcceptedCmp2, y=Store)) + 
    geom_boxplot(color="mediumorchid4", fill="mediumorchid1", alpha=0.2)
ggplot(new_data, aes(x=AcceptedCmp3, y=Store)) + 
    geom_boxplot(color="mediumorchid4", fill="mediumorchid1", alpha=0.2)
ggplot(new_data, aes(x=AcceptedCmp4, y=Store)) + 
    geom_boxplot(color="mediumorchid4", fill="mediumorchid1", alpha=0.2)
ggplot(new_data, aes(x=AcceptedCmp5, y=Store)) + 
    geom_boxplot(color="mediumorchid4", fill="mediumorchid1", alpha=0.2)
ggplot(new_data, aes(x=Response, y=Store)) + 
    geom_boxplot(color="mediumorchid4", fill="mediumorchid1", alpha=0.2)
```

People who accepted the campaigns 1, 2 and 5, have the highest median compared to others. Ergo they must have the highest store purchases.

```{r out.width=c('50%', '50%'), fig.show='hold'}
library(ggplot2)
ggplot(new_data, aes(x=AcceptedCmp1, y=WebVisitsMonth)) + 
    geom_boxplot(color="yellow3", fill="yellow", alpha=0.2)
ggplot(new_data, aes(x=AcceptedCmp2, y=WebVisitsMonth)) + 
    geom_boxplot(color="yellow3", fill="yellow", alpha=0.2)
ggplot(new_data, aes(x=AcceptedCmp3, y=WebVisitsMonth)) + 
    geom_boxplot(color="yellow3", fill="yellow", alpha=0.2)
ggplot(new_data, aes(x=AcceptedCmp4, y=WebVisitsMonth)) + 
    geom_boxplot(color="yellow3", fill="yellow", alpha=0.2)
ggplot(new_data, aes(x=AcceptedCmp5, y=WebVisitsMonth)) + 
    geom_boxplot(color="yellow3", fill="yellow", alpha=0.2)
ggplot(new_data, aes(x=Response, y=WebVisitsMonth)) + 
    geom_boxplot(color="yellow3", fill="yellow", alpha=0.2)
```

People who accepted the campaigns 1 and 5, have the lowest median compared to others. Ergo they must have the lowest rate of visits to website. This fact completely makes sense according to the above interpretation, because people who accepted these campaigns do their purchases in person and they don't buy online.

### The effect of campaigns on customers

#### Year_Birth:
First, we take a look at the distribution and details of the birth variable.

```{r}
new_data %>%
  ggplot( aes(x=Birth)) +
    geom_density(fill="blueviolet", color="aquamarine4", alpha=0.8) +
    ggtitle("distribution of Birth")
summary(Birth)
```

Now, in order to find out which age groups the advertising campaigns had an impact on, we use the following graphs.
Note that in each graph, the black color of the bars is a sign that the advertising campaign has no effect on that age group.

```{r}
table1=table(AcceptedCmp1,Birth)
prop.table=prop.table(table1,margin = 2)
barplot(prop.table,xlab = "AcceptedCmp1",ylab="",
        yaxt="n",main="distribution of AcceptedCmp1 BY Birth")
axis(2,at=(seq(0,1,0.2)),paste(seq(0,100,20),"%"))

table2=table(AcceptedCmp2,Birth)
prop.table=prop.table(table2,margin = 2)
barplot(prop.table,xlab = "AcceptedCmp2",ylab="",
        yaxt="n",main="distribution of AcceptedCmp2 BY Birth")
axis(2,at=(seq(0,1,0.2)),paste(seq(0,100,20),"%"))

table3=table(AcceptedCmp3,Birth)
prop.table=prop.table(table3,margin = 2)
barplot(prop.table,xlab = "AcceptedCmp3",ylab="",
        yaxt="n",main="distribution of AcceptedCmp3 BY Birth")
axis(2,at=(seq(0,1,0.2)),paste(seq(0,100,20),"%"))

table4=table(AcceptedCmp4,Birth)
prop.table=prop.table(table4,margin = 2)
barplot(prop.table,xlab = "AcceptedCmp4",ylab="",
        yaxt="n",main="distribution of AcceptedCmp4 BY Birth")
axis(2,at=(seq(0,1,0.2)),paste(seq(0,100,20),"%"))

table5=table(AcceptedCmp5,Birth)
prop.table=prop.table(table5,margin = 2)
barplot(prop.table,xlab = "AcceptedCmp5",ylab="",
        yaxt="n",main="distribution of AcceptedCmp5 BY Birth")
axis(2,at=(seq(0,1,0.2)),paste(seq(0,100,20),"%"))

table6=table(Response,Birth)
prop.table=prop.table(table6,margin = 2)
barplot(prop.table,xlab = "response",ylab="",
        yaxt="n",main="distribution of Response BY Birth")
axis(2,at=(seq(0,1,0.2)),paste(seq(0,100,20),"%"))
```

As it is clear in the graphs above, among different age groups, the latest campaign has been the most accepted by customers. The second campaign has the least probability of success in attracting customers.

#### Education:
Education has 5 levels. We want to investigate the effect of advertising campaigns on people with different education levels.

```{r out.width=c('50%', '50%'), fig.show='hold'}
library(RColorBrewer)
ed_acc1=table(Education,AcceptedCmp1)
col1=brewer.pal(5, "Reds") 
barplot(ed_acc1,
        legend.text = TRUE,
        beside = TRUE,
        main = "The impact of AcceptedCmp1",
        xlab = "AcceptedCmp1",
        ylab = "",col=col1)

ed_acc2=table(Education,AcceptedCmp2)
col2=brewer.pal(5, "Greens") 
barplot(ed_acc2,
        legend.text = TRUE,
        beside = TRUE,
        main = "The impact of AcceptedCmp2",
        xlab = "AcceptedCmp2",
        ylab = "",col=col2)

ed_acc3=table(Education,AcceptedCmp3)
col3=brewer.pal(5, "Blues") 
barplot(ed_acc3,
        legend.text = TRUE,
        beside = TRUE,
        main = "The impact of AcceptedCmp3",
        xlab = "AcceptedCmp3",
        ylab = "",col=col3)

ed_acc4=table(Education,AcceptedCmp4)
col4=brewer.pal(5, "Purples") 
barplot(ed_acc4,
        legend.text = TRUE,
        beside = TRUE,
        main = "The impact of AcceptedCmp4",
        xlab = "AcceptedCmp4",
        ylab = "",col=col4)

ed_acc5=table(Education,AcceptedCmp5)
col5=brewer.pal(5, "Greys") 
barplot(ed_acc5,
        legend.text = TRUE,
        beside = TRUE,
        main = "The impact of AcceptedCmp5",
        xlab = "AcceptedCmp5",
        ylab = "",col=col5)

ed_res=table(Education,Response)
col6=brewer.pal(5, "Oranges") 
barplot(ed_res,
        legend.text = TRUE,
        beside = TRUE,
        main = "The impact of Response",
        xlab = "Response",
        ylab = "",col=col6)
```

The rejection of all of the campaigns have almost the same distribution among different education levels.
The highest number of rejection among all education levels in all of the campaigns, is the graduation level. And the lowest of them all, is the basic level.

#### Marital Status:

```{r out.width=c('50%', '50%'), fig.show='hold'}
library(ggplot2)
ggplot(new_data, aes(x = Marital_Status,fill = AcceptedCmp1)) + geom_bar()
ggplot(new_data, aes(x = Marital_Status,fill = AcceptedCmp2)) + geom_bar()
ggplot(new_data, aes(x = Marital_Status,fill = AcceptedCmp3)) + geom_bar()
ggplot(new_data, aes(x = Marital_Status,fill = AcceptedCmp4)) + geom_bar()
ggplot(new_data, aes(x = Marital_Status,fill = AcceptedCmp5)) + geom_bar()
ggplot(new_data, aes(x = Marital_Status,fill = Response)) + geom_bar()
```

The highest acceptance rate in campaigns 1,3,4 and 5 belongs to married people. \\
And the above statement is true for campaign 2 with "together" status.
Above all, the last campaign had the most acceptance rate, especially in "single" status.

#### Income:

```{r}
library(ggplot2)
library(scales)
ggplot(new_data,aes(y = AcceptedCmp1,x = Income)) +
geom_jitter(alpha = 0.7,size = 1.5,color="mediumorchid4") + scale_x_continuous(label = dollar) +
labs(title = "Income by AcceptedCmp1",x = "Income",y = "AcceptedCmp1")

ggplot(new_data,aes(y = AcceptedCmp2,x = Income)) +
geom_jitter(alpha = 0.7,size = 1.5,color="seagreen3") + scale_x_continuous(label = dollar) +
labs(title = "Income by AcceptedCmp2",x = "Income",y = "AcceptedCmp2")

ggplot(new_data,aes(y = AcceptedCmp3,x = Income)) +
geom_jitter(alpha = 0.7,size = 1.5,color="turquoise4") + scale_x_continuous(label = dollar) +
labs(title = "Income by AcceptedCmp3",x = "Income",y = "AcceptedCmp3")

ggplot(new_data,aes(y = AcceptedCmp4,x = Income)) +
geom_jitter(alpha = 0.7,size = 1.5,color="mediumvioletred") + scale_x_continuous(label = dollar) +
labs(title = "Income by AcceptedCmp4",x = "Income",y = "AcceptedCmp4")

ggplot(new_data,aes(y = AcceptedCmp5,x = Income)) +
geom_jitter(alpha = 0.7,size = 1.5,color="orangered") + scale_x_continuous(label = dollar) +
labs(title = "Income by AcceptedCmp5",x = "Income",y = "AcceptedCmp5")

ggplot(new_data,aes(y = Response,x = Income)) +
geom_jitter(alpha = 0.7,size = 1.5,color="palegoldenrod") + scale_x_continuous(label = dollar) +
labs(title = "Income by Response",x = "Income",y = "Response")
```

Most of the people who accepted campaigns 1 and 5, had higher salary than others.
